\documentclass[master=mcs]{kulemt-tex/kulemt}
\setup{
  title={PIRANA implementation using GBFVtest},
  author={Ing. Antoine Janssens van der Maelen},
  promotor={Prof.\,Dr. Ir. F. Vercauteren},
  assessor={placeholder},
  assistant={Dr. Ir. R.~Geelen \and Dr. J.~Kang \and Ir. J.~Spiessens},}

% Remove the "%" on the next line for generating the cover page
% \setup{coverpageonly}
% Remove the "%" before the next "\setup" to generate only the first pages
% (e.g., if you are a Word user).
% \setup{frontpagesonly}

% If you want to include other LaTeX packages, do it here. 

% Finally the hyperref package is used for pdf files.
% This can be commented out for printed versions.
\usepackage[pdfusetitle,colorlinks,plainpages=false]{hyperref}
\usepackage{amsmath}
\usepackage{caption}
\usepackage{subcaption}%for 2 figures side by side
\usepackage{amssymb} 
\usepackage{enumitem}
\usepackage{msc} % Message Sequence Charts to make the protocle diagrams
\usepackage{tikz} % To draw figures
\usepackage{pgfplots} % For plotting data
\pgfplotsset{compat=1.18}
\usepackage{multirow} % For multirow in tables
% \usepackage{adjustbox} % To resize tables
\usepackage{amsthm} % For definition and theorems
\usepackage{algorithm} % For pseudocode
\usepackage{algpseudocode} % For pseudocode



%\includeonly{chapter-n}
\begin{document}
% Global paragraph spacing: no indent + extra vertical space
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.8em}

\begin{preface}
  First, I would like to express my gratitude to Dr. Ir. R. Geelen and Dr. J. Kang for their excellent supervision and for their guidance with the mathematical background. I would also like to thank Ir. J. Spiessens for creating EasyGBFV, which made working with GBFV much easier, and for his help in understanding the Fheanor library. Finally, I would like to thank KU Leuven, which introduced me to the fields of computer science and cryptography.
\end{preface}

\tableofcontents*

\begin{abstract}
  The \texttt{abstract} environment contains a more extensive overview of
  the work. But it should be limited to one page.

\end{abstract}

% A list of figures and tables is optional
%\listoffigures
%\listoftables
% If you only have a few figures and tables you can use the following instead
\listoffiguresandtables
% The list of symbols is also optional.
% This list must be created manually, e.g., as follows:
\chapter{List of Abbreviations and Symbols}
\input{chapters/xx_abbreviations}
\input{chapters/xx_symbols}

% Now comes the main text
\mainmatter

\include{chapters/1_intro}
\include{chapters/2_theoretical_background}
\include{chapters/3_PIR_implementation}
\include{chapters/4_results}
\include{chapters/5_discussion}
\include{chapters/x_conclusion}


% % If you have appendices:
% \appendixpage*          % if wanted
% \appendix
% \include{appendixes/appendix-1}
% % ... and so on until
% \include{appendixes/appendix-n}

\backmatter
% The bibliography comes after the appendices.
% You can replace the standard "abbrv" bibliography style by another one.
\bibliographystyle{abbrv}
\bibliography{references}

\end{document}
